---
title: "p8105_hw5_xm2356"
author: "Xinyin Miao (xm2356)"
date: "2025-11-11"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
set.seed(1)

knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.width = 7, fig.height = 4, dpi = 200
)


if (!dir.exists("data")) dir.create("data")
if (!dir.exists("results")) dir.create("results")
```

# Problem 1 

```{r}
bday_sim = function(n_room){
  birthdays = sample(1:365, n_room, replace = TRUE)
  
  repeated_bday = length(unique(birthdays)) < n_room
  
  repeated_bday
}


bday_sim_results = 
  expand_grid(
    bdays = 2:50,        
    iter = 1:10000        
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)   
  ) |> 
  group_by(
    bdays
  ) |> 
  summarize(
    prob_repeat = mean(result)
  )


bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Probability of at least one shared birthday",
    x = "Group size",
    y = "Estimated probability"
  ) +
  theme_minimal(base_size = 12)

```

The simulated results show a sharply increasing relationship between group size and the probability that at least two people share a birthday. When the group size is small (fewer than 10 people), the probability of a shared birthday is close to zero. As the number of people increases, this probability rises rapidly, reaching approximately 50% when there are about 23 individuals in the room. Beyond 40 people, the probability exceeds 90%, and by 50 it approaches certainty. These results highlight how counter-intuitive probability can be: even with only a few dozen people, coincidences such as shared birthdays become surprisingly likely. The smooth S-shaped curve confirms the robustness of the simulation and aligns closely with theoretical expectations.


# Problem 2

```{r}
n <- 30
sigma <- 5
alpha <- 0.05
mus <- 0:6
n_sims <- 5000

sim_one_mu <- function(mu, n_sims, n, sigma, alpha = 0.05) {
  tibble(sim = 1:n_sims) |>
    mutate(
      x = map(sim, ~ rnorm(n, mean = mu, sd = sigma)),
      t_out = map(x, ~ t.test(.x, mu = 0, conf.level = 1 - alpha)),
      t_tidy = map(t_out, broom::tidy)
    ) |>
    select(-x, -t_out) |>
    unnest(t_tidy) |>
    transmute(mu = mu, estimate = estimate, p_value = p.value)
}

```


```{r}
p2_sims <- map_dfr(mus, ~ sim_one_mu(.x, n_sims, n, sigma, alpha))

p2_power <- p2_sims |>
  group_by(mu) |>
  summarise(power = mean(p_value < alpha), .groups = "drop")


```


```{r}
p_power <- p2_power |>
  ggplot(aes(mu, power)) +
  geom_line() +
  geom_point(size = 1.5) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Power of one-sample t-test (n = 30, σ = 5, α = 0.05)",
    x = expression(true~mu),
    y = "Power (rejection rate)"
  ) +
  theme_minimal(base_size = 12)
```







